---
title: "Wrangling data (dplyr)"
subtitle: "CBio Thursday, session 3"
author: "Leon Eyrich Jessen & Johannes Eichler Waage"
date: "November 1st 2018"
output:
  ioslides_presentation:
    widescreen: true
    smaller: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library('tidyverse')
library('readxl')
```

## What is dplyr

dplyr is a grammar of data manipulation, providing a consistent set of verbs that help you solve the most common data manipulation challenges:

- `mutate()` adds new variables that are functions of existing variables
- `select()` picks variables based on their names
- `filter()` picks cases based on their values
- `summarise()` reduces multiple values down to a single summary
- `arrange()` changes the ordering of the rows

(For those in the room for whom it is relevant, there is a version called [`dbplyr`](https://cran.r-project.org/web/packages/dbplyr/vignettes/dbplyr.html), which can connect remote on-disk data stored in databases)

## Examples - data

The Danish Ministry of Economic and Business Affairs has made a publicly available data base called [noegletal.dk](http://www.noegletal.dk/). From here, we have extracted and cleaned [some example data](https://raw.githubusercontent.com/leonjessen/TidyThursday/master/data/kommune_data.tsv)

As we now know, we can read this directly of the web:

## Examples - data

```{r, echo=TRUE}
my_url = 'https://raw.githubusercontent.com/leonjessen/TidyThursday/master/data/kommune_data.tsv'
my_data = read_tsv(file = my_url)
```

## Examples - data

```{r, echo=TRUE}
my_data
```

As we can see, we have the fraction of citizens with a long education, the average tax-income for the municipality per citizen, how much money the municipality spends per primary school pupil and the fraction of pupils attending a private school.

## Examples - `mutate()`

`mutate()` creates new variables, lets calculate a new variable, which is the difference between how money is spend on each pupil and the average tax-income for the municipality per citizen:

```{r, echo=TRUE}
my_data = my_data %>% mutate(tax_in_pupil_exp_diff = Indt_indkskat - Folkeskudg_elev)
my_data
```

## Examples - `select()`

As we can see, our new variable is not shown, we can however use `select()` to show the new variable

```{r, echo=TRUE}
my_data %>% select(Kommune, Indt_indkskat, Folkeskudg_elev, tax_in_pupil_exp_diff)
```

## Examples - `filter()`

Let us find all the municipalities, who have a higher tax-income for the municipality per citizen, than what is spend on each pupil:

```{r, echo=TRUE}
my_data %>% filter(tax_in_pupil_exp_diff > 0)
```

The `filter()` function works on comparisons, i.e.

- Larger than `>`
- Smaller than `<`
- Equal to `==`
- Not equal to `!=`

## Comparisons

If you are new to comparisons, they are a little tricky initially. The result of a comparison is a logical, which can only ever be either `TRUE` or `FALSE`, e.g.

```{r, echo=TRUE}
2 > 3
2 < 3
2 == 3
2 != 3
```

## Comparisons

Comparisons can be combined using 'and' `&` (ampersand) 'or' `|` (pipe), where 'or' means "is either of these comparisons true?" 'and' means "are both of these comparisons true?", e.g.

```{r, echo=TRUE}
2 > 3 | 4 < 6
2 > 3 & 4 < 6
```

## Examples - `filter()`

Let us find all the municipalities, where less than 25% has a long education and where more than 25% of the pupils attend private school:

```{r, echo=TRUE}
my_data %>% filter(Videreg_udd <= 25 & Pct_prv_sk_elev >= 25)
```

## Examples - `summarise()`

We can calculate summary statistics using the `summary()` function:

```{r, echo=TRUE}
my_data %>% summarise(mu = mean(Folkeskudg_elev), sigma = sd(Folkeskudg_elev))
```

and if we combine this with the `group_by()` function, we can do this:

```{r, echo=TRUE}
my_data %>% group_by(Region) %>% summarise(mu = mean(Folkeskudg_elev), sigma = sd(Folkeskudg_elev))
```

## Examples - `arrange()`

This function simply arranges a tibble according to a varible, e.g. arrange ascending:

```{r, echo=TRUE}
my_data %>% arrange(Videreg_udd)
```

## Examples - `arrange()`

This function simply arranges a tibble according to a varible, e.g. arrange descending:

```{r, echo=TRUE}
my_data %>% arrange(desc(Videreg_udd))
```
