---
title: "Exercises: Data structures (tibble) and reading in data (readr)"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Ok, so now it is your turn!

Start by loading the tidyverse library (Type this command in the console in RStudio)

```{r, message=FALSE}
library('tidyverse')
```

### Question 1

R comes with some build in data sets, which can be used while learning R.

```{r}
diamonds
```

You can get information of the meaning of the variables in diamonds data set like so:

```{r, eval=FALSE}
?diamonds
```

Using one of the methods from the presentation, write the `diamonds` data set to an `excel_csv` file, then open it in excel and scroll down to row 27750.

__Q1__ What is the price of this diamond?

### Question 2

Now, again using one of the methods from the presentation, read in the `excel_csv` file you wrote to disk in question 1. We can use the `View()` function to look at data while staying in the RStudio IDE, like so:

```{r, eval=FALSE}
my_data %>% View
```

Using the `View()` function, look at the data and scroll down to row 27750

__Q2__ Is this the same value?

(For future reference, just do this `diamonds %>% slice(27750)`)

