---
title: "Data structures (tibble) and reading in data (readr)"
subtitle: "CBio Thursday, session 2"
author: "Leon Eyrich Jessen & Johannes Eichler Waage"
date: "November 1st 2018"
output:
  ioslides_presentation:
    widescreen: true
    smaller: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library('tidyverse')
library('readxl')
```

## Briefly on data types in R

In R the following data types exist

- Numbers are `numeric`, e.g. `c(1, 2, 3, 4)`
- Letters are `character` e.g. `c('A', 'B', 'C', 'D')`
- Yes/No are `logical`, e.g. `c(TRUE, FALSE, TRUE, FALSE)`
- Categories are `factor`, e.g. `factor(c('control', 'case', 'control', 'case'))`

We can check if a given variable is of a given data type, e.g.:

```{r, echo=TRUE}
is.numeric(c(1, 2, 3, 4))
```

## Briefly on Data structures in R

The matrix can only hold one data type, e.g. numerical
```{r base_matrix, echo=TRUE}
matrix(1:8, nrow = 2, byrow = TRUE)
```

The data frame can mix data types
```{r base_df, echo=TRUE}
(d = data.frame(x_var = 1:2, y_var = c('a','b')))
```

## Briefly on Data structures in R

Problem: `x_var` and  `y_var` can be accessed using `x` and `y`
```{r base_df_problem, echo=TRUE}
d$x
d$y
```

 - Short hand variable names
 - StringsAsFactors

# Let's be tidy

## Introducing the `tibble()`

Same data as before - but, now we no longer have a problem
```{r set_tibble_dat, echo=TRUE}
d = tibble(x_var = 1:2, y_var = c('a','b'))
```

```{r tibble_no_problem, echo=TRUE}
d$x
d$y_var
```

 - No short hand variable names
 - No stringsAsFactors

## Reading files

 - `read.table()` is now `read_table()`
 - `read.delim()` is now `read_delim()`
 - `read.delim(sep = "\t")` is now `read_tsv()`
 - `read.csv()` is now `read_csv()`
 - Etc.

In general, tidyverse uses the underscore for function names.

No need to use `fread()` the `tidyverse` read file functions are fast!

Also note we do not use one function with different delimiters, we use different functions for each file type! This improves code readability!

## Reading different types of files from disk

```{r, echo=TRUE, eval=FALSE}
my_data = read_excel(path = 'path/to/my/excel_file.xlsx', sheet = 'target_sheet')
my_data = read_csv(file = 'path/to/my/csv_file.csv')
my_data = read_tsv(file = 'path/to/my/tsv_file.tsv')
my_data = read_delim(file = 'path/to/my/delim_file.txt', delim = ';')
my_data = read_table(file = 'path/to/my/whitespace_file.txt')
```

## Reading different types of files from the web

```{r, echo=TRUE, eval=FALSE}
my_data = read_excel(path = 'http://www.my_site.com/path/to/files/excel_file.xlsx', sheet = 'target_sheet')
my_data = read_csv(file = 'http://www.my_site.com/path/to/files/csv_file.csv')
my_data = read_tsv(file = 'http://www.my_site.com/path/to/files/tsv_file.tsv')
my_data = read_delim(file = 'http://www.my_site.com/path/to/files/delim_file.txt', delim = ';')
my_data = read_table(file = 'http://www.my_site.com/path/to/files/whitespace_file.txt')
```

## Writing different types of files to disk

```{r, echo=TRUE, eval=FALSE}
write_csv(x = my_data, path = 'path/to/my/csv_file.csv')
write_tsv(x = my_data, path = 'path/to/my/tsv_file.tsv')
write_delim(x = my_data, path = 'path/to/my/delim_file.txt', delim = ';')
```

Note how `write_excel()` and `write_table` are missing. We do not export data to an excel file and we do not export data to mono-/multi whitespace separated files

## Other types of files

We can also read and write `SAS`/`SPSS`/`Stata` files using the [`haven`](https://github.com/tidyverse/haven) package

```{r, echo=TRUE, eval=FALSE}
install.packages('haven')
library('haven')

# SAS
read_sas("mtcars.sas7bdat")
write_sas(mtcars, "mtcars.sas7bdat")

# SPSS
read_sav("mtcars.sav")
write_sav(mtcars, "mtcars.sav")

# Stata
read_dta("mtcars.dta")
write_dta(mtcars, "mtcars.dta")
```

Or even write `Excel` files using the [`openxlsx`](https://readxl.tidyverse.org/#other-relevant-packages)

```{r, echo=TRUE, eval=FALSE}
install.packages('openxlsx')
library('openxlsx')
l <- list(iris = iris, mtcars = mtcars, chickwts = chickwts, quakes = quakes)
write.xlsx(l, file = "inst/extdata/datasets.xlsx")
```

## In summary

We hope, that it is evident, that you can pretty much interact with any other file format

Now, it is time for exercises!
